Single-fluorophore orientation determination with multiview polarized illumination: modeling and microscope design

### 2017-09-27 paper/paper/paper.tex
Change most instances of "molecule" to "dipole".

%%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
\begin{abstract*}
  We investigate the use of polarized illumination in multiview microscopes for
  determining the orientation of single fluorescent molecules. First, we relate
  the orientation of single molecules to measurable intensities in multiview
  microscopes and develop an information-theoretic metric---the solid-angle
  uncertainty---to compare the ability of multiview microscopes to estimate the
  orientation of single molecules. Next, we compare a broad class of microscopes
  using this metric---single- and dual-view microscopes with varying
  illumination polarization, illumination numerical aperture (NA), detection NA,
  obliquity, asymmetry, and exposure. We find that multiview microscopes are
  better at determining the orientation of single molecules than single-view
  microscopes. We also find that choosing a small illumination NA and a large
  detection NA are good design choices, that multiview microscopes can benefit
  from oblique illumination and detection, and that asymmetric NA microscopes
  can benefit from exposure asymmetry.
\end{abstract*}

@@

\section{Introduction}
The orientation of single molecules is a valuable reporter for
biological processes. By imaging biological samples with fluorescent probes that
are fixed to known structures, researchers have studied the dynamics of motor
proteins \cite{peterman2001, forkey2003}, DNA \cite{backer2016}, actin
\cite{mehta2016}, septin \cite{demay2011, mcquilken2017}, and membranes
\cite{anantharam2010}. The number of techniques available to biologists for
measuring the orientation of single molecules in live cells is continually
expanding. A recent review \cite{backlund2014} identified three major categories
of single-molecule orientation determination techniques:

\textbf{Spatial/angular variation of emission} techniques image single molecules
and use the distribution of intensity in the back focal plane \cite{lieb2004} or
in the image plane \cite{backer2014} to estimate their
orientation. Information-theoretically optimal techniques in this category use
defocusing or phase masks in the back focal plane to encode orientation and
position information in the image \cite{agrawal2012}. These techniques can
estimate the orientation of molecules in all orientations and only require a
single frame of intensity measurements, but they require complex reconstruction
algorithms, expensive high numerical aperture (NA) optics, and sensitive
detectors. Furthermore, these techniques spread the signal from a single
molecule over a large area on the detector, so they suffer from a poor
signal-to-noise ratio \cite{toprak2006}.

\textbf{Spatial variation of illumination polarization} techniques vary the
polarization and intensity in the illumination path to excite molecules in
specific orientations in the focal volume \cite{debarre2004}. Although these
techniques can accurately estimate the orientation of molecules in all
orientations, they require scanning which makes acquisition time a concern in
live-cell imaging.

\textbf{Polarized illumination or detection} techniques vary the polarization of
light in the illumination or detection paths and exploit the anisotropic
absorption and emission patterns of single molecules to estimate their
orientation. By illuminating and detecting from the focal plane, not just the
focal point, these techniques can estimate the orientation of many molecules in
parallel without scanning. These techniques are easy to implement---changing the
illumination or detection polarization is simple with a polarization splitter
\cite{mehta2016} or universal compensator/polarizer \cite{shribak2003}---and the
reconstruction methods are straightforward \cite{fourkas2001, mehta2016,
  backer2016}. Also, these techniques keep the signal from a single molecule
localized on the detector, so they have a signal-to-noise ratio advantage over
defocusing and phase masking techniques. The main drawback is that these
techniques require several intensity measurements under different polarization
orientations to estimate the orientation of a single molecule. Despite the
requirement for multiple measurements, polarized illumination or detection
methods are good choices for live-cell imaging because of their combination of
speed and ease of implementation.

@@

All single molecule orientation determination methods suffer from some degree of
anisotropic orientation uncertainty, i.e., some molecular orientations cannot be
determined as precisely as others. In some cases the orientation cannot be
determined at all---the forward model can be degenerate for specific dipole
orientations \cite{fourkas2001, lu2008}. We feel that the importance of
isotropic orientation uncertainty has been underappreciated. To our knowledge,
the only authors who have considered anisotropic orientation uncertainty in 3D
have only analyzed a subset of molecular orientations \cite{agrawal2012}. Our
view is that an ideal technique for determining the orientation of single
molecules can reconstruct the orientation with a small and nearly uniform
uncertainty for all molecular orientations.

@@

In this work we explore the use of polarized illumination in multiview
microscopes for determining the orientation of single molecules. Existing
multiview microscopes can easily be outfitted with fast-switching polarizers
that do not degrade image quality, so polarized illumination is a natural way to
augment multiview microscopes for measuring the orientation of single
molecules. Furthermore, multiview microscopes can achieve nearly isotropic
resolution while delivering selective illumination. We find that multiview
microscopes also provide a small and uniform orientation uncertainty for single
molecules in all orientations. We also find that single-view microscopes have
symmetries that can be broken by adding a second view. This reduction of
degeneracy allows multiview microscopes to uniquely determine the orientation of
single molecules.

@@

\section{Methods}\label{methods}
In this section we develop a method to compare multiview polarized illumination
microscopes for the task of reconstructing molecular orientations. In sections
\ref{excitation}--\ref{forward} we develop a model for the intensity collected
with a fixed illumination and detection configuration as a function of the
molecule orientation. In section \ref{designs} we list a variety of ways to
combine these intensity measurements into experimentally realizable polarized
illumination microscopes. Finally, in section \ref{metrics} we develop metrics
that we will use to compare these microscopes.

@@

\subsection{Absorption Efficiency}\label{excitation}
In this section we will calculate the \emph{absorption efficiency} of a single
molecule---the fraction of the incident power that is absorbed by the
molecule. Our approach is inspired by Fourkas \cite{fourkas2001}, but here we
calculate the absorption efficiency instead of the detection efficiency.

@@

  To find the absorption efficiency $\eta_{\text{abs}}$ of a single molecule if
  it were illuminated by a single ray, we take the unit incident electric field
  $\hat{\mb{p}}_{\text{exc}}$, take the dot product with the absorption dipole
  moment, then take the modulus squared
    \begin{align}
      \eta_{\text{abs}}^{\text{ray}} &= |\hat{\bs{\mu}}_{\text{abs}}\cdot\mh{p}_{\text{exc}}|^2. \label{eq:singleray}
    \end{align}

    To find the absorption efficiency of a single molecule, we integrate over
    all illumination rays and divide by the total incident power which gives the
    vector expression
\begin{align}
  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{p}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
\end{align}

We substitute equations \ref{eq:mu}--\ref{eq:rot_mat} into equation
\ref{eq:abs}, evaluate the integrals, and simplify to express the absorption
efficiency in scalar notation as
\begin{align}
  \eta_{\text{abs}} &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{exc}}})]\}\label{eq:scalarabs}
\end{align}
where
\begin{subequations}
\begin{align}
  A &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
  B &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
  C &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}\\
  D &= \frac{4}{3(1 - \cos\alpha)}.
\end{align}\label{eq:coefficients}%
\end{subequations}

We can extend these results to light-sheet illumination created by scanning a
laser. If we shine a narrow, collimated, laser beam onto a galvanometer in the
aperture plane and scan the galvanometer, we create a weakly focused light sheet
in the sample. If we scan the beam in the sample with a velocity much less than
the beam waist radius divided by the coherence time, then the coherence of the
laser is rendered ineffectual by the scanning---the light sheet is effectively
composed of many incoherent beams. If we also ensure that the laser beam is
weakly focused, we can ignore the excitation caused by the small fields aligned
with the optical axis. Under slowly-scanned and weakly-focused light-sheet
illumination, molecules in the sample will be excited as if a single plane wave
was incident. Therefore, we can find the absorption efficiency of a molecule in
this case by taking the limit of equation \ref{eq:scalarabs} as
$\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
\ref{eq:excfield} into equation \ref{eq:singleray}) giving
\begin{align}
  \eta_{\text{abs}}^{\text{ray}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}}),
\end{align}
which is recognizable as Malus' law generalized to three dimensions. This means
that light-sheet illumination is approximately equivalent to low-NA illumination
with respect to the molecular orientations they excite. Notice that the
absorption efficiency can take its maximum range of values
($\eta_{\text{abs}} \in [0, 1]$) when $\alpha=0$.

\subsection{Detection Efficiency}\label{detection}
In this section we will calculate the \emph{detection efficiency}---the fraction
of the power emitted by a single molecule that we detect. Fourkas calculated the
detection efficiency when an ideal objective with its optical axis aligned with
the $\mh{z}$ axis is focused on a single molecule and a polarizer is placed in
the aperture plane \cite{fourkas2001}. He found that
\begin{align}
  \eta_{\text{det}}^{\text{pol}} = A + B\sin^2\Theta + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{det}}})], \label{eq:det_pol}
\end{align}
where $\phi_{\text{det}}$ is the angle between the transmission axis of the
detection polarizer and the positive $\mh{x}$ axis.

@@

\subsection{Intensity Measurements}\label{forward}
The detected intensity is proportional to the product of the sample irradiance
$I_{\text{in}}$, the absorption efficiency, and the detection efficiency
\begin{align}
  I \propto I_{\text{in}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame},
\end{align}
where the proportionality depends on the exposure time, the molecular quantum
efficiencies, and other constant factors. Equation \ref{eq:forward_frame} is the
forward model for a single intensity measurement. Figure \ref{fig:single-frame}
shows the efficiencies in three representative geometries. Note that in all
cases there are multiple orientations that give the same intensity
measurement. This means that a single intensity measurement does not give us
enough information to reconstruct the orientation of a single molecule. Next, we
will consider combining several intensity measurements to create polarized
illumination microscopes that can reconstruct the orientation of single
molecules.

@@

\subsection{Evaluation Metrics}\label{metrics}
Our goal is to evaluate the ability of a microscope design to estimate the
orientation of a single molecule (via parameters $\Theta$ and $\Phi$) from
intensity data. A common way to evaluate the ability to estimate the parameters
$\Theta$ and $\Phi$ from the data is to calculate the Cram\'er-Rao lower bound
(CRLB) for each parameter \cite{kay1993}. The CRLBs are given by the diagonal
elements of the inverse of the Fisher information matrix, and they give the
minimum variance of an unbiased estimator for each parameter. For each
microscope design and molecule orientation we calculate the Fisher information
matrix. We assume that the detected intensities are Poisson distributed, so the
Fisher information matrix is given by
\begin{align}
  \mb{F} &= \sum_{k=1}^N \frac{1}{I_k}
  \begin{bmatrix}
    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Theta}&\dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}\\\\
    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}&\dfrac{\partial I_k}{\partial \Phi}\dfrac{\partial I_k}{\partial \Phi}\\    
  \end{bmatrix}
\end{align}
where $I_k$ is the $k$th intensity measurement in a microscope that uses $N$
intensity measurements. Agrawal et. al. used the square root of the product of
the CRLBs multiplied by the Jacobian determinant,
$\sin\Theta\sqrt{\mb{F}^{-1}_{0,0}\mb{F}^{-1}_{1,1}}$, to find the area of
uncertainty in parameter space \cite{agrawal2012}.

@@

For each microscope we calculate $\sigma_{\Omega}$ at 10,000 approximately
equally spaced points on the unit sphere. A desirable microscope design will
have a small solid-angle uncertainty that is uniform for all molecular
orientations. The most straightforward way to find the location and scale of the
solid-angle uncertainty is to use the mean and variance, respectively. However,
the solid-angle uncertainty can take extremely large values, e.g., when a
fluorescent dipole is aligned with the optical axis of a single excitation
objective. Such large values can change the mean and variance dramatically
depending on the sample points, so we use the median and median absolute
deviation (MAD)---the median of the data's absolute difference from the
median---as robust alternatives to the mean and variance.

@@

We also use the Fisher information matrix to calculate uncertainty ellipses for
each microscope and fluorophore orientation. The general equation of an ellipse
with axis radii $a$ and $b$ and rotation angle $\gamma$ is
\begin{align}
  \frac{(x\cos\gamma + y\sin\gamma)^2}{a^2} + \frac{(x\cos\gamma - y\sin\gamma)^2}{b^2} =1. 
\end{align}
We calculate the ellipse parameters in terms of the Fisher information matrix
using the following set of equations
\begin{align}
  \sigma_x &= \sqrt{\mb{F}^{-1}_{0,0}} \label{eq:first}\\
  \sigma_y &= \sin\Theta\sqrt{\mb{F}^{-1}_{1,1}} \label{eq:second}\\ 
  \sigma_{xy} &= \sin\Theta\sqrt{\mb{F}^{-1}_{1,0}} \label{eq:third}\\  
  a^2 &= \frac{\sigma_x^2 + \sigma_y^2}{2} + \sqrt{\frac{(\sigma_x^2 - \sigma_y^2)^2}{4} + \sigma_{xy}^2}\\
  b^2 &= \frac{\sigma_x^2 + \sigma_y^2}{2} - \sqrt{\frac{(\sigma_x^2 - \sigma_y^2)^2}{4} + \sigma_{xy}^2}\\
  \gamma &= \frac{1}{2}\text{arctan}\left(\frac{2\sigma_{xy}}{\sigma_x^2 - \sigma_y^2}\right). \label{eq:last}
\end{align}
Equations \ref{eq:first}--\ref{eq:last} are identical to those found in
\cite{coe2009} with an additional factor of $\sin\Theta$ on equations
\ref{eq:second} and \ref{eq:third} to account for the non-Euclidean geometry of
the sphere. Instead of plotting ellipses with widely varying radii on the same
figure, we chose a constant major axis radius. This means that the uncertainty
ellipses indicate the relative size of uncertainty in different directions, not
the absolute size of the uncertainty.

The solid-angle uncertainty and the uncertainty ellipse are good local measures
of orientation uncertainty, but they are insensitive to degeneracies with
distant dipole orientations. For example, all of the intensity measurements in
this paper are identical under inversion of the dipole orientation
($I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in equations \ref{eq:scalarabs},
\ref{eq:scalardet}, and \ref{eq:forward_frame}), but the solid-angle uncertainty
and the uncertainty ellipse give no indication of this degeneracy. For this
reason we also use the number of degeneracies as a measure of a microscope's
ability to estimate the orientation of a dipole from intensity measurements. A
small number of degeneracies is important for a microscope to find the
orientation of a molecule---a twofold degenerate microscope can find the
orientation of a molecule down to the hemisphere while a fourfold degenerate
microscope can only find the orientation down to the quadrant.

@@

\section{Results}\label{results}
Figure \ref{fig:single-arm} shows our results for single-view designs where a
single objective is used for illumination and detection. We swept through the
illumination and detection NA while keeping the sample irradiance constant, and
we found that the lowest median and MAD of the solid-angle uncertainty occurs
with a small illumination NA and a large detection NA where
$\alpha \approx 65^{\circ}$ (see Figure \ref{fig:single-arm}e) at
NA${}_{\text{det}} \approx 1.2$). A small illumination NA maximizes the range of
absorption efficiencies, while a large detection NA optimizes the number of
detected photons and the range of the detection efficiencies. The optimal
detection NA is slightly larger than $\alpha = 57.4^{\circ}$ because increasing
the detection NA reduces the effect of shot noise. Note the relative importance
of illumination and detection NAs---reducing the illumination NA (an inexpensive
modification requiring underfilling the aperture plane) improves the solid-angle
uncertainty much less than increasing the detection NA (an expensive
modification requiring a higher-NA objective). Also note in Figure
\ref{fig:single-arm}c) that single-view designs suffer from high orientation
uncertainty when the molecules are oriented along the optical axis (the
molecules are not efficiently excited or detected), near the transverse plane
(it is ambiguous whether the molecule is oriented above or below the transverse
plane, see Figure \ref{fig:single-arm}b)), and near the polarizer orientations
(it is ambiguous whether the molecule is on either side of the polarizer
orientation, see Figure \ref{fig:single-arm}b)).

>>>>

%%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
\begin{abstract*}
  We investigate the use of polarized illumination in multiview microscopes for
  determining the orientation of single-molecule fluorescence transition
  dipoles. First, we relate the orientation of single dipoles to measurable
  intensities in multiview microscopes and develop an information-theoretic
  metric---the solid-angle uncertainty---to compare the ability of multiview
  microscopes to estimate the orientation of single dipoles. Next, we compare
  a broad class of microscopes using this metric---single- and dual-view
  microscopes with varying illumination polarization, illumination numerical
  aperture (NA), detection NA, obliquity, asymmetry, and exposure. We find that
  multi-view microscopes can measure all dipole orientations, while the
  orientations measurable with single-view microscopes is halved because of
  symmetries in the detection process. We also find that choosing a small
  illumination NA and a large detection NA are good design choices, that
  multiview microscopes can benefit from oblique illumination and detection, and
  that asymmetric NA microscopes can benefit from exposure asymmetry.
\end{abstract*}

@@

\section{Introduction}
The orientation of single-molecule fluorescence transition dipoles is a valuable
reporter for biological processes. By imaging biological samples with
fluorescent probes that are fixed to known structures, researchers have studied
the dynamics of motor proteins \cite{peterman2001, forkey2003}, DNA
\cite{backer2016}, actin \cite{mehta2016}, septin \cite{demay2011,
  mcquilken2017}, and membranes \cite{anantharam2010}. The number of techniques
available to biologists for measuring the orientation of single dipoles in live
cells is continually expanding. A recent review \cite{backlund2014} identified
three major categories of single-dipole orientation determination techniques:

\textbf{Spatial/angular variation of emission} techniques image single molecules
and use the distribution of intensity in the back focal plane \cite{lieb2004} or
in the image plane \cite{backer2014} to estimate their dipole
orientation. Information-theoretically optimal techniques in this category use
defocusing or phase masks in the back focal plane to encode orientation and
position information in the image \cite{agrawal2012}. These techniques can
estimate the orientation of dipoles in all orientations and only require a
single frame of intensity measurements, but they require complex reconstruction
algorithms, expensive high numerical aperture (NA) optics, and sensitive
detectors. Furthermore, these techniques spread the signal from a single
molecule over a large area on the detector, so they suffer from a poor
signal-to-noise ratio \cite{toprak2006}.

\textbf{Spatial variation of illumination polarization} techniques vary the
polarization and intensity in the illumination path to excite dipoles in
specific orientations in the focal volume \cite{debarre2004}. Although these
techniques can accurately estimate the orientation of dipoles in all
orientations, they require scanning which makes acquisition time a concern in
live-cell imaging.

\textbf{Polarized illumination or detection} techniques vary the polarization of
light in the illumination or detection paths and exploit the anisotropic
absorption and emission patterns of single molecules to estimate their dipole
orientation. By illuminating and detecting from the focal plane, not just the
focal point, these techniques can estimate the orientation of many dipoles in
parallel without scanning. These techniques are easy to implement---changing the
illumination or detection polarization is simple with a polarization splitter
\cite{mehta2016} or universal compensator/polarizer \cite{shribak2003}---and the
reconstruction methods are straightforward \cite{fourkas2001, mehta2016,
  backer2016}. Also, these techniques keep the signal from a single molecule
localized on the detector, so they have a signal-to-noise ratio advantage over
defocusing and phase masking techniques. The main drawback is that these
techniques require several intensity measurements under different polarization
orientations to estimate the orientation of a single dipole. Despite the
requirement for multiple measurements, polarized illumination or detection
methods are good choices for live-cell imaging because of their combination of
speed and ease of implementation.

@@

All single-dipole orientation determination methods suffer from some degree of
anisotropic orientation uncertainty, i.e., some dipole orientations cannot be
determined as precisely as others. In some cases the orientation cannot be
determined at all---the forward model can be degenerate for specific dipole
orientations \cite{fourkas2001, lu2008}. We feel that the importance of
isotropic orientation uncertainty has been underappreciated. To our knowledge,
the only authors who have considered anisotropic orientation uncertainty in 3D
have only analyzed a subset of dipole orientations \cite{agrawal2012}. Our view
is that an ideal technique for determining the orientation of single dipoles can
reconstruct the orientation with a small and nearly uniform uncertainty for all
dipole orientations.

@@

In this work we explore the use of polarized illumination in multiview
microscopes for determining the orientation of single dipoles. Existing
multiview microscopes can easily be outfitted with fast-switching polarizers
that do not degrade image quality, so polarized illumination is a natural way to
augment multiview microscopes for measuring the orientation of single
dipoles. Furthermore, multiview microscopes can achieve nearly isotropic
resolution while delivering selective illumination. We find that multiview
microscopes also provide a small and uniform orientation uncertainty for single
dipoles in all orientations. We also find that single-view microscopes have
symmetries that can be broken by adding a second view. This reduction of
degeneracy allows multiview microscopes to uniquely determine the orientation of
single dipoles.

@@

\section{Methods}\label{methods}
In this section we develop a method to compare multiview polarized illumination
microscopes for the task of reconstructing dipole orientations. In sections
\ref{excitation}--\ref{forward} we develop a model for the intensity collected
with a fixed illumination and detection configuration as a function of the
dipole orientation. In section \ref{designs} we list a variety of ways to
combine these intensity measurements into experimentally realizable polarized
illumination microscopes. Finally, in section \ref{metrics} we develop metrics
that we will use to compare these microscopes.

@@

\subsection{Absorption Efficiency}\label{excitation}
In this section we will calculate the \emph{absorption efficiency} of a single
dipole---the fraction of the incident power that is absorbed by the
dipole. Our approach is inspired by Fourkas \cite{fourkas2001}, but here we
calculate the absorption efficiency instead of the detection efficiency.

@@

  To find the absorption efficiency $\eta_{\text{abs}}$ of a single dipole if
  it were illuminated by a single ray, we take the unit incident electric field
  $\hat{\mb{p}}_{\text{exc}}$, take the dot product with the absorption dipole
  moment, then take the modulus squared
    \begin{align}
      \eta_{\text{abs}}^{\text{ray}} &= |\hat{\bs{\mu}}_{\text{abs}}\cdot\mh{p}_{\text{exc}}|^2. \label{eq:singleray}
    \end{align}

    To find the absorption efficiency of a single dipole, we integrate over
    all illumination rays and divide by the total incident power which gives the
    vector expression
\begin{align}
  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{p}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
\end{align}

We substitute equations \ref{eq:mu}--\ref{eq:rot_mat} into equation
\ref{eq:abs}, evaluate the integrals, and simplify to express the absorption
efficiency in scalar notation as
\begin{align}
  \eta_{\text{abs}} &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{exc}}})]\}\label{eq:scalarabs}
\end{align}
where
\begin{subequations}
\begin{align}
  A &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
  B &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
  C &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}\\
  D &= \frac{4}{3(1 - \cos\alpha)}.
\end{align}\label{eq:coefficients}%
\end{subequations}
Note that the factor $D$ keeps the specimen irradiance constant for any value of
$\alpha$.

We can extend these results to light-sheet illumination created by scanning a
laser. If we shine a narrow, collimated, laser beam onto a galvanometer in the
aperture plane and scan the galvanometer, we create a weakly focused light sheet
in the sample. If we scan the beam in the sample with a velocity much less than
the beam waist radius divided by the coherence time, then the coherence of the
laser is rendered ineffectual by the scanning---the light sheet is effectively
composed of many incoherent beams. If we also ensure that the laser beam is
weakly focused, we can ignore the excitation caused by the small fields aligned
with the optical axis. Under slowly-scanned and weakly-focused light-sheet
illumination, dipoles in the sample will be excited as if a single plane wave
was incident. Therefore, we can find the absorption efficiency of a dipole in
this case by taking the limit of equation \ref{eq:scalarabs} as
$\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
\ref{eq:excfield} into equation \ref{eq:singleray}) giving
\begin{align}
  \eta_{\text{abs}}^{\text{ray}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}}),
\end{align}
which is recognizable as Malus' law generalized to three dimensions. This means
that light-sheet illumination is approximately equivalent to low-NA illumination
with respect to the dipole orientations they excite. Notice that the
absorption efficiency can take its maximum range of values
($\eta_{\text{abs}} \in [0, 1]$) when $\alpha=0$.

\subsection{Detection Efficiency}\label{detection}
In this section we will calculate the \emph{detection efficiency}---the fraction
of the power emitted by a single dipole that we detect. Fourkas calculated the
detection efficiency when an ideal objective with its optical axis aligned with
the $\mh{z}$ axis is focused on a single dipole and a polarizer is placed in
the aperture plane \cite{fourkas2001}. He found that
\begin{align}
  \eta_{\text{det}}^{\text{pol}} = A + B\sin^2\Theta + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{det}}})], \label{eq:det_pol}
\end{align}
where $\phi_{\text{det}}$ is the angle between the transmission axis of the
detection polarizer and the positive $\mh{x}$ axis.

@@

\subsection{Intensity Measurements}\label{forward}
The detected intensity is proportional to the product of the sample irradiance
$I_{\text{in}}$, the absorption efficiency, and the detection efficiency
\begin{align}
  I \propto I_{\text{in}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame},
\end{align}
where the proportionality depends on the exposure time, the molecular quantum
efficiencies, and other constant factors. Equation \ref{eq:forward_frame} is the
forward model for a single intensity measurement. Figure \ref{fig:single-frame}
shows the efficiencies in three representative geometries. Note that in all
cases there are multiple orientations that give the same intensity
measurement. This means that a single intensity measurement does not give us
enough information to reconstruct the orientation of a single dipole. Next, we
will consider combining several intensity measurements to create polarized
illumination microscopes that can reconstruct the orientation of single
dipoles.

@@

\subsection{Evaluation Metrics}\label{metrics}
Our goal is to evaluate the ability of a microscope design to estimate the
orientation of a single dipole (via parameters $\Theta$ and $\Phi$) from
intensity data. A common way to evaluate the ability to estimate the parameters
$\Theta$ and $\Phi$ from the data is to calculate the Cram\'er-Rao lower bound
(CRLB) for each parameter \cite{kay1993}. The CRLBs are given by the diagonal
elements of the inverse of the Fisher information matrix, and they give the
minimum variance of an unbiased estimator for each parameter. For each
microscope design and dipole orientation we calculate the Fisher information
matrix. We assume that the detected intensities are Poisson distributed, so the
Fisher information matrix is given by
\begin{align}
  \mb{F} &= \sum_{k=1}^N \frac{1}{I_k}
  \begin{bmatrix}
    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Theta}&\dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}\\\\
    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}&\dfrac{\partial I_k}{\partial \Phi}\dfrac{\partial I_k}{\partial \Phi}\\    
  \end{bmatrix}
\end{align}
where $I_k$ is the $k$th intensity measurement in a microscope that uses $N$
intensity measurements. Agrawal et. al. used the square root of the product of
the CRLBs multiplied by the Jacobian determinant,
$\sin\Theta\sqrt{\mb{F}^{-1}_{0,0}\mb{F}^{-1}_{1,1}}$, to find the area of
uncertainty in parameter space \cite{agrawal2012}.

@@

For each microscope we calculate $\sigma_{\Omega}$ at 10,000 approximately
equally spaced points on the unit sphere. A desirable microscope design will
have a small solid-angle uncertainty that is uniform for all dipole
orientations. The most straightforward way to find the location and scale of the
solid-angle uncertainty is to use the mean and variance, respectively. However,
the solid-angle uncertainty can take extremely large values, e.g., when a
fluorescent dipole is aligned with the optical axis of a single excitation
objective. Such large values can change the mean and variance dramatically
depending on the sample points, so we use the median and median absolute
deviation (MAD)---the median of the data's absolute difference from the
median---as robust alternatives to the mean and variance.

@@

We also use the Fisher information matrix to calculate uncertainty ellipses for
each microscope and dipole orientation. The general equation of an ellipse
with axis radii $a$ and $b$ and rotation angle $\gamma$ is
\begin{align}
  \frac{(x\cos\gamma + y\sin\gamma)^2}{a^2} + \frac{(x\cos\gamma - y\sin\gamma)^2}{b^2} =1. 
\end{align}
We calculate the ellipse parameters in terms of the Fisher information matrix
using the following set of equations
\begin{align}
  \sigma_x &= \sqrt{\mb{F}^{-1}_{0,0}} \label{eq:first}\\
  \sigma_y &= \sin\Theta\sqrt{\mb{F}^{-1}_{1,1}} \label{eq:second}\\ 
  \sigma_{xy} &= \sin\Theta\sqrt{\mb{F}^{-1}_{1,0}} \label{eq:third}\\  
  a^2 &= \frac{\sigma_x^2 + \sigma_y^2}{2} + \sqrt{\frac{(\sigma_x^2 - \sigma_y^2)^2}{4} + \sigma_{xy}^2}\\
  b^2 &= \frac{\sigma_x^2 + \sigma_y^2}{2} - \sqrt{\frac{(\sigma_x^2 - \sigma_y^2)^2}{4} + \sigma_{xy}^2}\\
  \gamma &= \frac{1}{2}\text{arctan}\left(\frac{2\sigma_{xy}}{\sigma_x^2 - \sigma_y^2}\right). \label{eq:last}
\end{align}
Equations \ref{eq:first}--\ref{eq:last} are identical to those found in
\cite{coe2009} with an additional factor of $\sin\Theta$ on equations
\ref{eq:second} and \ref{eq:third} to account for the non-Euclidean geometry of
the sphere. Instead of plotting ellipses with widely varying radii on the same
figure, we chose a constant major axis radius. This means that the uncertainty
ellipses indicate the relative size of uncertainty in different directions, not
the absolute size of the uncertainty.

The solid-angle uncertainty and the uncertainty ellipse are good local measures
of orientation uncertainty, but they are insensitive to degeneracies with
distant dipole orientations. For example, all of the intensity measurements in
this paper are identical under inversion of the dipole orientation
($I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in equations \ref{eq:scalarabs},
\ref{eq:scalardet}, and \ref{eq:forward_frame}), but the solid-angle uncertainty
and the uncertainty ellipse give no indication of this degeneracy. For this
reason we also use the number of degeneracies as a measure of a microscope's
ability to estimate the orientation of a dipole from intensity measurements. A
small number of degeneracies is important for a microscope to find the
orientation of a dipole---a twofold degenerate microscope can find the
orientation of a dipole down to the hemisphere while a fourfold degenerate
microscope can only find the orientation down to the quadrant.

@@

\section{Results}\label{results}
Figure \ref{fig:single-arm} shows our results for single-view designs where a
single objective is used for illumination and detection. We swept through the
illumination and detection NA while keeping the sample irradiance constant, and
we found that the lowest median and MAD of the solid-angle uncertainty occurs
with a small illumination NA and a large detection NA where
$\alpha \approx 65^{\circ}$ (see Figure \ref{fig:single-arm}e) at
NA${}_{\text{det}} \approx 1.2$). A small illumination NA maximizes the range of
absorption efficiencies, while a large detection NA optimizes the number of
detected photons and the range of the detection efficiencies. The optimal
detection NA is slightly larger than $\alpha = 57.4^{\circ}$ because increasing
the detection NA reduces the effect of shot noise. Note the relative importance
of illumination and detection NAs---reducing the illumination NA (an inexpensive
modification requiring underfilling the aperture plane) improves the solid-angle
uncertainty much less than increasing the detection NA (an expensive
modification requiring a higher-NA objective). Also note in Figure
\ref{fig:single-arm}c) that single-view designs suffer from high orientation
uncertainty when the dipoles are oriented along the optical axis (the
dipoles are not efficiently excited or detected), near the transverse plane
(it is ambiguous whether the dipole is oriented above or below the transverse
plane, see Figure \ref{fig:single-arm}b)), and near the polarizer orientations
(it is ambiguous whether the dipole is on either side of the polarizer
orientation, see Figure \ref{fig:single-arm}b)).

The single-view microscope in Figure \ref{fig:single-arm} is fourfold degenerate
($I(\Theta, \Phi) = I(\Theta + \pi, \Phi) = I(\Theta, \Phi + \pi) = I(\Theta +
\pi, \Phi+\pi)$ in equations \ref{eq:scalarabs}, \ref{eq:scalardet}, and
\ref{eq:forward_frame}). The first twofold degeneracy is due to the usual
inversion degeneracy---the dipole is an axis rather than a direction. A second
twofold degeneracy is present in single-view designs---the microscope cannot
distinguish between dipoles reflected through the transverse plane. Evidence of
the second degeneracy is found in Figures \ref{fig:single-arm}b) and
\ref{fig:single-arm}c) which show the ambiguity of whether the dipole is
oriented above or below the transverse plane. Near the transverse plane this
ambiguity is evident in the high solid-angle uncertainty, but far from the
transverse plane evidence of the ambiguity is lost. Adding a second view and
taking eight intensity measurements instead of four removes this degeneracy (see
equations \ref{eq:thetap} and \ref{eq:phip}), so the dual-view microscopes we
consider in this paper are only twofold degenerate. This means that while
dual-view microscopes can find the orientation of a dipole up to the
hemisphere, single-view microscopes can only find the orientation of a dipole
up to the quadrant.
